---
title: "Hands-on Exercise 4"
author: "Oh Jia Wen"
date: 5/4/2023
date-modified: 5/4/2023
execute: 
  warning: false
---

# Getting started

1.  Using p_load() of pacman package to load the required libraries

```{r}
pacman::p_load(ggstatsplot, tidyverse,rstantools)
```

2.  Importing data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

# Plotting the graph

## 1) Visual Statistical Analysis

### 1.1) One-sample test: **gghistostats()** method

In the below output, `gghistostats()` is used to to build an visual of one-sample test on English scores.

#### Output:

```{r}
set.seed(1234)

gghistostats(
  data = exam_data,
  x = ENGLISH,
  type = "bayes",
  test.value = 60,
  xlab = "English scores"
)
```

Default information: - statistical details - Bayes Factor - sample sizes - distribution summary

A Bayes factor is the ratio of the likelihood of one particular hypothesis to the likelihood of another. It can be interpreted as a measure of the strength of evidence in favor of one theory among two competing theories.

### 1.2) Two-sample mean test: **ggbetweenstats()**

In the below output, `ggbetweenstats()` is used to to build an visual of two-sample mean test.

#### Output:

```{r}
ggbetweenstats(
  data = exam_data,
  x = GENDER, 
  y = MATHS,
  type = "np",
  messages = FALSE
)
```

Default information: - statistical details - Bayes Factor - sample sizes - distribution summary

### 1.3) One-way ANOVA Test: **ggbetweenstats()** method

In the below output, `ggbetweenstats()` is used to to build an visual of one-way ANOVA test.

#### Output:

```{r}
ggbetweenstats(
  data = exam_data, 
  x = RACE, 
  y = ENGLISH, 
  type = "p", 
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE,
  #"ns" for only non-significant, "s" for only significant, "all" for everything
  pairwise.display = "s",      
  p.adjust.method = "fdr", 
  messages = FALSE 
  )
```

-   "ns" → only non-significant

-   "s" → only significant

-   "all" → everything

### 1.4) Significant Test of Correlation: **ggscatterstats()**

In the below output, `ggscatterstats()` is used to to build a visual for Significant Test of Correlation between Maths scores and English scores.

#### Output:

```{r}
ggscatterstats(
  data = exam_data,
  x = MATHS,
  y = ENGLISH,
  marginal = FALSE,
  )
```

### 1.5) Significant Test of Association (Depedence) : **ggbarstats()** methods

In the below output, `ggbarstats()` is used to to build a visual for Significant Test of Association (Dependence).

In the code chunk below, the Maths scores is binned into a 4-class variable by using [cut().]{.underline}

#### Output:

```{r}
exam1 <- exam_data |> 
  mutate(MATHS_bins =
           cut(MATHS, 
               breaks = c(0, 60, 75, 85, 100)))
```

In this code chunk below, `ggbarstats()` is used to build a visual for Significant Test of Association.

```{r}
ggbarstats(
  data = exam1,
  x = MATHS_bins,
  y = GENDER
)
```

## 2) Visualising Models

### 2.1) Getting Started

1.  Using p_load() of pacman package to load the required libraries

```{r}
pacman::p_load(readxl, performance, parameters, see)
```

2.  Importing data

```{r}
car_resale <- read_xls("data/ToyotaCorolla.xls", 
                       "data")
car_resale
```

Note: `car_resale` is a tibble data frame.

### 2.2) Multiple Regression Model using **lm()**

In the below output, `lm()` is used to calibrate a multiple linear regression model of Base Stats of R.

#### Output:

```{r}
model <- lm(Price ~ Age_08_04 + Mfg_Year + KM + 
              Weight + Guarantee_Period, data = car_resale)
model
```

### 2.3) Model Diagnostic: Checking for multicolinearity using **check_collinearity()**

In the below output, `check_collinearity()` is used to check for multicolinearity.

#### Output:

```{r}
check_collinearity(model)
```

```{r}
check_c <- check_collinearity(model)
plot(check_c)
```

As per figure above, Age_08_04 and Mfg_Year are highly correlated.

Remove/Drop Mfg_Year.

### 2.3) Model Diagnostic: Checking for normality assumption using **check_normality()**

In the below output, `check_normality()` is used to check for normality assumption. \#### Output:

```{r}
model1 <- lm(Price ~ Age_08_04 + KM + 
              Weight + Guarantee_Period, data = car_resale)
```

```{r}
check_n <- check_normality(model1)
```

```{r}
plot(check_n)
```

Based on the figure above, we will reject the Null hypothesis and infer that the model failed to conform to normaility assumption.

### 2.4) Model Diagnostic: Checking model for homogeneity of variances using **check_heteroscedasticity()**

In the below output, `check_heteroscedasticity()` is used to check model for homogeneity of variances.

#### Output:

```{r}
check_h <- check_heteroscedasticity(model1)
```

```{r}
plot(check_h)
```

### 2.5) Model Diagnostic: Complete check using **check_model()**

In the below output, `check_model()` is used.

#### Output:

```{r}
check_model(model1)
```

### 2.6) Visualising Regression Parameters

In the below output, `plot()` of see package and `parameters()` of parameters package are used to visualize the parameters of a regression model.

#### Output:

```{r}
plot(parameters(model1))
```

In the below output, `ggcoefstats()` of ggstatsplot package is used to visualize the parameters of a regression model.

#### Output:

```{r}
ggcoefstats(model1, 
            output = "plot")
```

## 3) Visualising Uncertainty

### 3.1) Getting Started

1.  Using p_load() of pacman package to load the required libraries

```{r}
pacman::p_load(tidyverse, plotly, crosstalk, DT, ggdist, gganimate)
```

2.  Importing data

```{r}
exam <- read_csv("data/Exam_data.csv")
```

### 3.2) Visualizing the uncertainty of point estimates of using **ggplot2**

In the below output, `ggplot2` is used.

It performs group by RACE and compute by calculating the observation, mean, standard deviation, and standard error of MATHS score by RACE. Then, it saves the output as a tibble data table called my_sum.

#### Output:

```{r}
my_sum <- exam_data |> 
  group_by(RACE) |> 
  summarize(
    n = n(),
    mean = mean(MATHS),
    sd = sd(MATHS)) |>
  mutate(se = sd/sqrt(n-1))
my_sum$RACE  <- fct_reorder(my_sum$RACE, my_sum$mean, .desc = TRUE)
```

Next, the code chunk below will be shown in html format.

```{r}
knitr::kable(head(my_sum), format = 'html')
```

#### 3.2.1 ) Reveal the standard error of mean Maths score by race

```{r}
ggplot(my_sum) +
  geom_errorbar(
    aes(x=RACE, 
        ymin=mean-se, 
        ymax=mean+se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    size=0.5) +
  geom_point(aes
           (x=RACE, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  ggtitle("Standard error of mean 
          maths score by rac")
```

#### 3.2.2 ) Reveal the 95% confidence interval of mean Maths score by race

Error bars sorted by the average maths scores.

```{r}
ggplot(my_sum) +
  geom_errorbar(
    aes(x=RACE, 
        ymin=mean - 1.96*se, 
        ymax=mean + 1.96*se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    size=0.5) +
  geom_point(aes
           (x=RACE, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  ggtitle(" 95% confidence interval of mean maths score by race ")
```

#### 3.2.3) Visualizing the uncertainty of point estimates with interactive error bars

```{r}
d <- highlight_key(my_sum) 

p <- ggplot(my_sum) +
  geom_errorbar(
    aes(x=RACE, 
        ymin=mean - 2.58*se, 
        ymax=mean + 2.58*se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    size=0.5) +
  
    geom_point(aes
           (x=RACE, 
            y=mean,
            text = paste("Race:", RACE,
                     "<br>N:", n,
                     "<br>Avg. Scores:", round(mean, digits = 2),
                     "<br>99% CI:[", round(mean - 2.58*se, digits = 2), ", ", round(mean + 2.58*se, digits = 2), "]")), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  
  ggtitle(" 99% confidence interval of mean maths score by race ")

gg <- highlight(ggplotly(p,tooltip = "text"),        
                "plotly_selected")  

dt <- DT::datatable(d,
                    colnames = c(" ", "No. of pupils", "Avg Scores", "Std Dev", "Std Error")) |> 
  formatRound(columns = c("mean", "sd", "se"), digits = 2)

crosstalk::bscols(gg,               
                  dt, 
                  widths = 5)        
```

### 3.3) Visualizing Uncertainty using **ggdist**

ggdist is an R package that provides a flexible set of ggplot2 geoms and stats designed especially for visualizing distributions and uncertainty.

It is designed for both frequentist and Bayesian uncertainty visualization, taking the view that uncertainty visualization can be unified through the perspective of distribution visualization:

-   for frequentist models, one visualizes confidence distributions or bootstrap distributions (see vignette("freq-uncertainty-vis"));

-   for Bayesian models, one visualizes probability distributions (see the tidybayes package, which builds on top of ggdist).

#### 3.3.1 ) Using **stat_pointinterval()** of ggdist

In the code chunk below, stat_pointinterval() of ggdist is used to build a visual for displaying distribution of maths scores by race.

```{r}
exam %>%
  ggplot(aes(x = RACE, 
             y = MATHS)) +
  stat_pointinterval() +   #<<
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Mean Point + Multiple-interval plot")
```

```{r}
exam %>%
  ggplot(aes(x = RACE, y = MATHS)) +
  stat_pointinterval(.width = 0.95,
  .point = median,
  .interval = qi) +
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Mean Point + Multiple-interval plot")
```

#### 3.3.2 ) Visualization of 95% and 99% confidence interval with mean

```{r}
exam %>%
  ggplot(aes(x = RACE, 
             y = MATHS)) +
  stat_pointinterval(
  .point = mean,
  .interval = c(qi(0.95),qi(0.99))) +   
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Mean Point + Multiple-interval plot")
```

#### 3.3.3 ) Visualizing the uncertainty of point estimates: **ggdist** methods

Using `stat_gradientinterval()` of ggdist is used to build a visual for displaying distribution of maths scores by race.

```{r}
exam %>%
  ggplot(aes(x = RACE, 
             y = MATHS)) +
  stat_gradientinterval(   
    fill = "skyblue",      
    show.legend = TRUE     
  ) +                        
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Gradient + interval plot")
```

### 3.4) Visualising Uncertainty with **Hypothetical Outcome Plots (HOPs)**

#### 3.4.1) Getting Started

1.  Installing ungeviz package

```{r}
devtools::install_github("wilkelab/ungeviz")
```

2.  Launch the application in R

```{r}
library(ungeviz)
```

```{r}
ggplot(data = exam, 
       (aes(x = factor(RACE), y = MATHS))) +
  geom_point(position = position_jitter(
    height = 0.3, width = 0.05), 
    size = 0.4, color = "#0072B2", alpha = 1/2) +
  geom_hpline(data = sampler(25, group = RACE), height = 0.6, color = "#D55E00") +
  theme_bw() + 
  # `.draw` is a generated column indicating the sample draw
  transition_states(.draw, 1, 3)
```

## 4) Funnel Plots with R

Funnel plot is a specially designed data visualisation for conducting unbiased comparison between outlets, stores or business entities.

### 4.1) Getting Started

1.  Using p_load() of pacman package to load the required libraries

-   Five R packages will be used. They are:

    -   **readr** for importing csv into R.

    -   **FunnelPlotR** for creating funnel plot.

    -   **ggplot2** for creating funnel plot manually.

    -   **knitr** for building static html table.

    -   **plotly** for creating interactive funnel plot.

```{r}
pacman::p_load(tidyverse, FunnelPlotR, plotly, knitr)
```

2.  Importing data

```{r}
covid19 <- read_csv("data/COVID-19_DKI_Jakarta.csv") %>%
  mutate_if(is.character, as.factor)
```

### 4.2) FunnelPlotR methods

FunnelPlotR package uses ggplot to generate funnel plots. It requires a numerator (events of interest), denominator (population to be considered) and group. The key arguments selected for customization are:

-   limit: plot limits (95 or 99).

-   label_outliers: to label outliers (true or false).

-   Poisson_limits: to add Poisson limits to the plot.

-   OD_adjust: to add overdispersed limits to the plot. xrange and yrange: to specify the range to display for axes, acts like a zoom function.

-   Other aesthetic components such as graph title, axis labels etc.

#### 4.2.1) FunnelPlotR methods: **The basic plot**

```{r}
funnel_plot(
  numerator = covid19$Positive,
  denominator = covid19$Death,
  group = covid19$`Sub-district`
)
```
A funnel plot object with 267 points of which 0 are outliers. 
Plot is adjusted for overdispersion. 

##### 4.2.1.1) Makeover 1

```{r}
funnel_plot(
  numerator = covid19$Death,
  denominator = covid19$Positive,
  group = covid19$`Sub-district`,
  data_type = "PR",     #<<
  xrange = c(0, 6500),  #<<
  yrange = c(0, 0.05)   #<<
)
```
A funnel plot object with 267 points of which 7 are outliers. 
Plot is adjusted for overdispersion. 

##### 4.2.1.2) Makeover 2

```{r}
funnel_plot(
  numerator = covid19$Death,
  denominator = covid19$Positive,
  group = covid19$`Sub-district`,
  data_type = "PR",   
  xrange = c(0, 6500),  
  yrange = c(0, 0.05),
  label = NA,
  title = "Cumulative COVID-19 Fatality Rate by Cumulative Total Number of COVID-19 Positive Cases", #<<           
  x_label = "Cumulative COVID-19 Positive Cases", #<<
  y_label = "Cumulative Fatality Rate"  #<<
)
```
A funnel plot object with 267 points of which 7 are outliers. 
Plot is adjusted for overdispersion. 

### 4.3) FunnelPlotR methods: **ggplot2** methods 

To begin with, we derive cumulative death rate and standard error of cumulative death rate.

**Data Preparation**

```{r}
df <- covid19 %>%
  mutate(rate = Death / Positive) %>%
  mutate(rate.se = sqrt((rate*(1-rate)) / (Positive))) %>%
  filter(rate > 0)
```

Next, the fit.mean is computed by using the code chunk below.

```{r}
fit.mean <- weighted.mean(df$rate, 1/df$rate.se^2)
```

**Calculate lower and upper limits for 95% and 99.9% CI**

```{r}
number.seq <- seq(1, max(df$Positive), 1)
number.ll95 <- fit.mean - 1.96 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ul95 <- fit.mean + 1.96 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ll999 <- fit.mean - 3.29 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ul999 <- fit.mean + 3.29 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
dfCI <- data.frame(number.ll95, number.ul95, number.ll999, 
                   number.ul999, number.seq, fit.mean)
```

**Plotting a static funnel plot**

```{r}
p <- ggplot(df, aes(x = Positive, y = rate)) +
  geom_point(aes(label=`Sub-district`), 
             alpha=0.4) +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ll95), 
            size = 0.4, 
            colour = "grey40", 
            linetype = "dashed") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ul95), 
            size = 0.4, 
            colour = "grey40", 
            linetype = "dashed") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ll999), 
            size = 0.4, 
            colour = "grey40") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ul999), 
            size = 0.4, 
            colour = "grey40") +
  geom_hline(data = dfCI, 
             aes(yintercept = fit.mean), 
             size = 0.4, 
             colour = "grey40") +
  coord_cartesian(ylim=c(0,0.05)) +
  annotate("text", x = 1, y = -0.13, label = "95%", size = 3, colour = "grey40") + 
  annotate("text", x = 4.5, y = -0.18, label = "99%", size = 3, colour = "grey40") + 
  ggtitle("Cumulative Fatality Rate by Cumulative Number of COVID-19 Cases") +
  xlab("Cumulative Number of COVID-19 Cases") + 
  ylab("Cumulative Fatality Rate") +
  theme_light() +
  theme(plot.title = element_text(size=12),
        legend.position = c(0.91,0.85), 
        legend.title = element_text(size=7),
        legend.text = element_text(size=7),
        legend.background = element_rect(colour = "grey60", linetype = "dotted"),
        legend.key.height = unit(0.3, "cm"))
p
```

**Interactive Funnel Plot: plotly + ggplot2**

The funnel plot created using ggplot2 functions can be made interactive with `ggplotly()` of plotly r package.

```{r}
fp_ggplotly <- ggplotly(p,
  tooltip = c("label", 
              "x", 
              "y"))
fp_ggplotly
```


