---
title: "Hands-on Exercise 6 Part Two"
author: "Oh Jia Wen"
date: 5/20/2023
date-modified: 5/20/2023
execute: 
  echo: true
  eval: true
  warning: false
---

# Getting started

1.  Using p_load() of pacman package to load the required libraries

```{r}
pacman::p_load(corrplot, ggstatsplot, seriation, dendextend, heatmaply, GGally, parallelPlot, tidyverse)
```

2.  Importing Data

```{r}
wine <- read_csv("data/wine_quality.csv")
```

## 1. Plotting Correlation Matrix: pairs() method

Create a scatterplot matrix by using the pairs function of R Graphics.

### 1.1 Basic Correlation Matrix

```{r}
pairs(wine[,1:11])
```

Another example. This time, it uses column 2 to 12 to the build the scatter plot matrix.

```{r}
pairs(wine[,2:12])
```

### 1.2 Lower Half of Correlation Matrix

Since correlation matrix is symmetric, we would like to see the lower half or upper half. Thus, we would need to choose between upper.panel = NULL or lower.panel = NULL .

```{r}
pairs(wine[,2:12], upper.panel = NULL)
```

### 1.3 Show Correlation Coefficient

panel.cor function will be used to show the correlation coefficient of each pair of variables instead of a scatter plot.

```{r}
```

## 2. Plotting Correlation Matrix: ggcormat()

We will be using ggcorrmat() of ggstatsplot package.
