---
title: "Hands-on Exercise 6 Part Two"
author: "Oh Jia Wen"
date: 5/20/2023
date-modified: 5/20/2023
execute: 
  echo: true
  eval: true
  warning: false
---

# Getting started

1.  Using p_load() of pacman package to load the required libraries

```{r}
pacman::p_load(corrplot, ggstatsplot, tidyverse)
```

2.  Importing Data

```{r}
wine <- read_csv("data/wine_quality.csv")
```

## 1. Plotting Correlation Matrix: pairs() method

Create a scatterplot matrix by using the pairs function of R Graphics.

### 1.1 Basic Correlation Matrix

```{r}
pairs(wine[,1:11])
```

Another example. This time, it uses column 2 to 12 to the build the scatter plot matrix.

```{r}
pairs(wine[,2:12])
```

### 1.2 Lower Half of Correlation Matrix

Since correlation matrix is symmetric, we would like to see the lower half or upper half. Thus, we would need to choose between upper.panel = NULL or lower.panel = NULL .

```{r}
pairs(wine[,2:12], upper.panel = NULL)
```

### 1.3 Show Correlation Coefficient

panel.cor function will be used to show the correlation coefficient of each pair of variables instead of a scatter plot.

```{r}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...) {
usr <- par("usr")
on.exit(par(usr))
par(usr = c(0, 1, 0, 1))
r <- abs(cor(x, y, use="complete.obs"))
txt <- format(c(r, 0.123456789), digits=digits)[1]
txt <- paste(prefix, txt, sep="")
if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
text(0.5, 0.5, txt, cex = cex.cor * (1 + r) / 2)
}

pairs(wine[,2:12], 
      upper.panel = panel.cor)
```

## 2. Plotting Correlation Matrix: ggcormat()

We will be using ggcorrmat() of ggstatsplot package.

```{r}
#|fig.height: 12
#|fig.width: 6

ggstatsplot::ggcorrmat(
  data = wine, 
  cor.vars = 1:11,
  ggcorrplot.args = list(outline.color = "black", 
                         hc.order = TRUE,
                         tl.cex = 10),
  title    = "Correlogram for wine dataset",
  subtitle = "Four pairs are no significant at p < 0.05"
)
```

```{r}
#help to control specific component of the plot such 
ggplot.component = list(
    theme(text=element_text(size=5),
      axis.text.x = element_text(size = 8),
      axis.text.y = element_text(size = 8)))
```

### 2. Plotting Correlation Matrix: ggcormat()

`ggstatsplot` package supports faceting. However the feature is not available in *ggcorrmat()* but in the [*grouped_ggcorrmat()*](https://indrajeetpatil.github.io/ggstatsplot/reference/grouped_ggcorrmat.html) of **ggstatsplot**.

```{r}
#|fig.height: 12
#|fig.width: 6
grouped_ggcorrmat(
  data = wine,
  cor.vars = 1:11,
  grouping.var = type,        #to build facet plot
  type = "robust",
  p.adjust.method = "holm",
  
  #provides list of additional arguments
  plotgrid.args = list(ncol = 2),       
  ggcorrplot.args = list(outline.color = "black", 
                         hc.order = TRUE,
                         tl.cex = 10),
  
  #calling plot annotations arguments of patchwork
  annotation.args = list(               
    tag_levels = "a",
    title = "Correlogram for wine dataset",
    subtitle = "The measures are: alcohol, sulphates, fixed acidity, citric acid, chlorides, residual sugar, density, free sulfur dioxide and volatile acidity",
    caption = "Dataset: UCI Machine Learning Repository"
  )
)
```

## 3. Plotting Corrplot package: corrplot()

We will be using corrplot() of the corrplot package. Check out -\> [An Introduction to corrplot Package](https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html) for basic understanding of corrplot package.

```{r}
wine.cor <- cor(wine[, 1:11])
```

Next, [*corrplot()*](https://www.rdocumentation.org/packages/corrplot/versions/0.2-0/topics/corrplot) is used to plot the corrgram by using all the default setting as shown in the code chunk below.

```{r}
corrplot(wine.cor)
```

### 3.1 Further Visualization with Corrplot()

There are seven visual geometrics (parameter method) can be used to encode the attribute values. They are: circle, square, ellipse, number, shade, color and pie. The default is circle.

#### Example 1: 

```{r}
corrplot(wine.cor, 
         method = "ellipse",  #default is circle 
         type="lower",        #default is full
         diag = FALSE,        #turn off diagonal celss
         tl.col = "black")    #change axis text label color to black 
```

#### Example 2: 

```{r}
corrplot.mixed(wine.cor, 
               lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               tl.col = "black")
```

#### Example 3: Combine corrgram with significant test 

```{r}
wine.sig = cor.mtest(wine.cor, conf.level= .95)
```

```{r}
corrplot(wine.cor,
         method = "number",
         type = "lower",
         diag = FALSE,
         tl.col = "black",
         tl.srt = 45,
         p.mat = wine.sig$p,  #input the calculated conf.level
         sig.level = .05)
```

#### Reordering corrgram: 

Matrix reorder is very important for mining the hiden structure and pattern in a corrgram. By default, the order of attributes of a corrgram is sorted according to the correlation matrix (i.e. \"original\"). The default setting can be over-write by using the *order* argument of *corrplot()*. Currently, **corrplot** package support four sorting methods, they are:

-   \"AOE\" is for the angular order of the eigenvectors. See Michael Friendly (2002) for details.

-   \"FPC\" for the first principal component order.

-   \"hclust\" for hierarchical clustering order, and \"hclust.method\" for the agglomeration method to be used.

    -   \"hclust.method\" should be one of \"ward\", \"single\", \"complete\", \"average\", \"mcquitty\", \"median\" or \"centroid\".

-   \"alphabet\" for alphabetical order.

\"AOE\", \"FPC\", \"hclust\", \"alphabet\". More algorithms can be found in **seriation** package.

```{r}
corrplot.mixed(wine.cor, 
               lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",        #instead of diag, change to order = "hclust"
               order="AOE",   #order_type.  then hcluster.method = "ward.D"
               tl.col = "black")
```

## 4. Plotting Heatmap

1.  Using p_load() of pacman package to load the required libraries

```{r}
pacman::p_load(seriation, dendextend, heatmaply, tidyverse)
```

2.  Importing data

```{r}
wh <- read_csv("data/WHData-2018.csv")
```

3.  Preparing the data

```{r}
row.names(wh) <- wh$Country
```

4.  Transform to Matrix

```{r}
wh1 <- dplyr::select(wh, c(3, 7:12))
wh_matrix <- data.matrix(wh)
```

### 4.1 heatmap() of R Stats

We will be using *heatmap()* of Base Stats.

```{r}
wh_heatmap <- heatmap(wh_matrix,
                      Rowv=NA, Colv=NA)
```

To plot cluster heatmap,

```{r}
wh_heatmap <- heatmap(wh_matrix)
```

### 4.2 Interactive Heatmap

We will be using heatmaply in this section.

**heatmaply** supports a variety of hierarchical clustering algorithm. The main arguments provided are:

-   *distfun*: function used to compute the distance (dissimilarity) between both rows and columns. Defaults to dist. The options \"pearson\", \"spearman\" and \"kendall\" can be used to use correlation-based clustering, which uses as.dist(1 - cor(t(x))) as the distance metric (using the specified correlation method).

-   *hclustfun*: function used to compute the hierarchical clustering when *Rowv* or *Colv* are not dendrograms. Defaults to *hclust*.

-   *dist_method* default is NULL, which results in \"euclidean\" to be used. It can accept alternative character strings indicating the method to be passed to distfun. By default *distfun* is \"dist\"\" hence this can be one of \"euclidean\", \"maximum\", \"manhattan\", \"canberra\", \"binary\" or \"minkowski\".

-   *hclust_method* default is NULL, which results in \"complete\" method to be used. It can accept alternative character strings indicating the method to be passed to *hclustfun*. By default hclustfun is hclust hence this can be one of \"ward.D\", \"ward.D2\", \"single\", \"complete\", \"average\" (= UPGMA), \"mcquitty\" (= WPGMA), \"median\" (= WPGMC) or \"centroid\" (= UPGMC).

#### Clustering: Manual Approach 

In the code chunk below, the heatmap is plotted by using hierachical clustering algorithm with \"Euclidean distance\" and \"ward.D\" method.

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "ward.D")
```

#### Clustering: Statistical Approach 

In order to determine the best clustering method and number of cluster the *dend_expend()* and *find_k()* functions of **dendextend** package will be used.

First, the *dend_expend()* will be used to determine the recommended clustering method to be used.

```{r}
wh_d <- dist(normalize(wh_matrix[, -c(1, 2, 4, 5)]), method = "euclidean")
dend_expend(wh_d)[[3]]
```

The output table shows that \"average\" method should be used because it gave the high optimum value.

Next, *find_k()* is used to determine the optimal number of cluster.

```{r}
wh_clust <- hclust(wh_d, method = "average")
num_k <- find_k(wh_clust)
plot(num_k)
```

Figure above shows that k=3 would be good.

With reference to the statistical analysis results, we can prepare the code chunk as shown below.

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "average",
          k_row = 3)
```

#### Seriation:

**heatmaply** uses the seriation package to find an optimal ordering of rows and columns.

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "OLO")
```

The default options is \"OLO\" (Optimal leaf ordering) which optimizes the above criterion (in O(n\^4)). Another option is \"GW\" (Gruvaeus and Wainer) which aims for the same goal but uses a potentially faster heuristic.

#### Final outcome: 

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          Colv=NA,
          seriate = "none",
          colors = Blues,
          k_row = 5,
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 5,
          main="World Happiness Score and Variables by Country, 2018 \nDataTransformation using Normalise Method",
          xlab = "World Happiness Indicators",
          ylab = "World Countries"
          )
```
