---
title: "Take Home Exercise 1"
subtitle: " Visual Analytics of the demographic and financial characteristics of residents in City of Engagement"
author: "Oh Jia Wen"
date: "5/13/2023"
date-modified: "5/13/2023"
execute: 
  echo: true
  eval: true
  warning: false
---

# 1. OVERVIEW

City of Engagement is a small city located at Country of Nowhere, with a total population of 50,000, serving as an agriculture region. The local council of the city is in the midst of preparing the Local Plan 2023.

## 1.1 The Task

In this take-home exercise, patterns containing demographics and financial characteristics of residents in City of Engagement will be unveiled using appropriate static and interactive statistical graphics methods.

# 2. Datasets

Data has been collected by the local council of the city. The survey sampled 1,000 respondents to collect data related to their household demographic, spending patterns and among, other things. The data is stored in two separate files ranging from **Mar 2022 to Feb 2023** :

|           | 1\) Participants.csv | 2\) FinancialJourval.csv |
|:---------:|:--------------------:|:------------------------:|
|   Rows    |        1,011         |        1,513,636         |
| Variables |          7           |            4             |

## 2.1 Metadata

+----------------------+----------------+-----------------------------------------------------------------------------+
| File                 | Variables Name | Description                                                                 |
+:====================:+:==============:+:===========================================================================:+
| participants.csv     | participantId  | Unique identification to represent the Participants                         |
|                      |                |                                                                             |
| FinancialJourval.csv |                |                                                                             |
+----------------------+----------------+-----------------------------------------------------------------------------+
| participants.csv     | householdSize  | Represents the number of people in the household                            |
+----------------------+----------------+-----------------------------------------------------------------------------+
| participants.csv     | haveKids       | Binary value (True/False) indicating if participant have a kid(s)           |
+----------------------+----------------+-----------------------------------------------------------------------------+
| participants.csv     | age            | Represents age of the participant                                           |
+----------------------+----------------+-----------------------------------------------------------------------------+
| participants.csv     | educationLevel | Represents the highest education attained by participant                    |
+----------------------+----------------+-----------------------------------------------------------------------------+
| FinancialJourval.csv | interestGroup  | Represents the group associated with the participant                        |
+----------------------+----------------+-----------------------------------------------------------------------------+
| FinancialJourval.csv | joviality      | Represents the level of happiness by participant at the start of the survey |
+----------------------+----------------+-----------------------------------------------------------------------------+
| FinancialJourval.csv | timestamp      | Represents the date and time the entry was inputted                         |
+----------------------+----------------+-----------------------------------------------------------------------------+
| FinancialJourval.csv | category       | Represents the type of income/expenses incurred at a given timestamp        |
+----------------------+----------------+-----------------------------------------------------------------------------+
| FinancialJourval.csv | amount         | Represents the amount received ( *+ income),* amount paid *( - expenses )*  |
+----------------------+----------------+-----------------------------------------------------------------------------+

# 3. Data Preparation

## 3.1 Install R-packages

Using p_load() of pacman package to load and install the following libraries:

-   `ggiraph` : For creating interactive 'ggplot' graphics

-   `plotly` : For creating interactive statistical graphs

-   `patchwork` : For combining multiple ggplot2 graphs into one figure

-   `tidyverse` : A collection of R packages use in everyday data analyses. It is able to support data science, data wrangling, and analysis.

-   `knitr`: For dynamic report generation

-   `ggstatsplot`: For creating graphics with details from statistical tests included and its plot

-   `paletteer`: Collection of color palettes

-   `wesanderson`: *Wes Anderson*'s theme Palette Generator

```{r}
#to double check if got use DT, scales, ggpubr
pacman::p_load(ggiraph, plotly, patchwork, DT, tidyverse,
               knitr,scales,ggstatsplot,paletteer,wesanderson,
               ggpubr,crosstalk,gganimate, ggridges, rstatix,gt,webshot2,png,nortest)
options(scipen = 999)
```

::: callout-tip
## Tip

`options(scipen = 999)` : removes scientific notation in our exercise.
:::

## 3.2 Import Data

### 3.2.1 Import participants dataset

```{r}
participants <- read_csv("data/Participants.csv")
```

### 3.2.2 Load participants

::: panel-tabset
## Data Table

```{r}
#| echo: false
head(participants)
```

## Code

```{r}
#| eval: false
head(participants)
```
:::

### 3.2.3 Import Financial Journal dataset

```{r}
financial_journal <- read_csv("data/FinancialJournal.csv")
```

### 3.2.4 Load Financial Journal

::: panel-tabset
#### Data Table

```{r}
#| echo: false
head(financial_journal)
```

#### Code

```{r}
#| eval: false
head(financial_journal)
```
:::

## 3.3 Data Wrangling

As seen from the two data tables above, there are some issues that could be rectify. Henceforth, the following adjustments are made:

**3.3.1 participants.csv :**

1.  participantId is a `<dbl>` variable. (Rectify by reformatting it to `<chr>`)

2.  householdSize is a `<dbl>` variable. (Revised to `<ord>` for the order of categories)

3.  age is a continuous variable which makes it harder to visualize the demographics (Create a new column with 5-class variables after determining the youngest and oldest demographics.)

```{r}
#| code-fold: true
#check min and max age of residents in COE. 
min(participants$age)
max(participants$age)
```

4.  educationLevel is a `<chr>` variable. (Revised to `<ord>` for the order of categories)

5.  joviality has nine decimal places. (Rectify by rounding it to 2.d.p and create new 5-class variables for future analysis)

```{r}
#| code-fold: true
#create new dataset 
participants_new <- participants %>%
    mutate(
          participantId = as.character(participantId),
          #binned joviality to 5-class variables 
          joviality_bins = cut(joviality, breaks = c(0.0,0.2,0.4,0.6,0.8,1.0))
          )

#reformat householdSize to Ordinal 
    participants_new$householdSize <- factor(participants$householdSize,
                                      levels = c("1", "2", "3"), 
                                      ordered = TRUE) 
#reformat age group 
participants_new$age_group <- factor(ifelse(participants$age < 20, "Under 20",
                ifelse(participants$age < 30, "20-29",
                    ifelse(participants$age < 40, "30-39",
                      ifelse(participants$age < 50, "40-49", "Above 50")))),
                levels = c("Under 20", "20-29", "30-39", "40-49", "Above 50"),
                ordered= TRUE)

#reformat education level to Ordinal 
participants_new$educationLevel <- factor(participants$educationLevel, 
                                      levels = c("Low", "HighSchoolOrCollege", 
                                                 "Bachelors", "Graduate"
                                                 ), 
                                      ordered = TRUE)  

#round up joviality to 2 decimal places 
participants_new$joviality <- round(participants$joviality, 2) 

#output data frame
participants_new

```

**3.3.2 FinancialJourval.csv :**

1.  Similar issue as point 1 above.

2.  Timestamp is a `<POSIX>` variable. (Rectify by reformatting it to `<chr>` in *year-mth* format)

3.  As per the code below, there are duplicate entries in the financial journal. (Rectify by using the `distinct()` function from `[dplyr package]` )

```{r}
#| code-fold: true
#check for duplicates 
dup <- (nrow(financial_journal) - nrow(unique(financial_journal)))
#reformat output 
dup_reformat <- format(dup, big.mark=",")
#print output
dup_reformat
```

4.  "Category" is not very useful. (Rectify by using `pivot_wider()` function from `[tidyr package]` to transpose)

```{r}
#| code-fold: true

    #remove duplicate rows for all columns
    financial_journal_lessdup <- financial_journal %>% 
      distinct()

    #create new dataset 
    grouped_data <- financial_journal_lessdup %>%
      
    #recode participantId from dbl to chr, format timestamp to year_mth and round amount 
        mutate(participantId = as.character(participantId),
             year_mth = format(as.Date(financial_journal_lessdup$timestamp), "%Y-%m"),
             amount = abs(round(amount,2)),
             .before = 3) %>%
      
    #group the columns in the following order and sum the amount to total_amount
      group_by( participantId, year_mth,category) %>%
      summarize(total_amount = sum(amount)) 

    # pivot the data frame to have categories as columns
    pivoted_fj <- grouped_data %>%
      pivot_wider(names_from = "category", 
                  values_from = "total_amount", values_fill = 0)

    # create new column from list of categories
    pivoted_fj$Shelter_new <- pivoted_fj$Shelter + pivoted_fj$RentAdjustment
    pivoted_fj$Expenses <- pivoted_fj$Education + pivoted_fj$Food +  
    pivoted_fj$Recreation + pivoted_fj$Shelter + pivoted_fj$RentAdjustment
    pivoted_fj$Income <- pivoted_fj$Wage
    pivoted_fj$Cashflow <- pivoted_fj$Income - pivoted_fj$Expenses

    # output the pivoted data frame
    pivoted_fj %>%
      select(c(1:5,7,9:12))
```

5.  Type of Income/Expenses are all labelled in a column. (Create new columns)

    +----------------------+----------------+------------------------------------------------------------------------------------+
    | Data Table           | Variables Name | Description                                                                        |
    +:====================:+:==============:+:==================================================================================:+
    | resident_profile_rev | Income         | Category :Wage                                                                     |
    +----------------------+----------------+------------------------------------------------------------------------------------+
    | resident_profile_rev | Expenses       | Category: Education + Recreation + Food + Shelter_new *(Shelter + RentAdjustment)* |
    +----------------------+----------------+------------------------------------------------------------------------------------+
    | resident_profile_rev | Cashflow       | NEW : Income - Expenses                                                            |
    +----------------------+----------------+------------------------------------------------------------------------------------+

6.  Multiple zero values for RentAdjustment. As seen from the data frame above and the code chunk below, there are only 72 rows. Thus, it will be combined with Shelter.

```{r}
colSums(pivoted_fj[-1] !=0)
```

## 3.4 Merging of Data frame

`Full_join` will be used to create a new table by joining the cleaned participants file and pivoted financial journal. It will be match by participant's ID. Likewise, the sequence in the dataset will be relocated to highlight several columns.

```{r}
#| code-fold: true

#join both data sets 
resident_profile <- full_join(participants_new, pivoted_fj, 
                       by = c("participantId" = "participantId")) %>%
#relocate columns to the front (by importance)
                    relocate(year_mth, .after =participantId) %>%
                    relocate(Cashflow, .after = year_mth) %>%
                    relocate(age_group, .after = Cashflow) %>%
                    relocate(educationLevel, .after = age_group) %>%
                    relocate(Income, .after = haveKids) %>%
                    relocate(Expenses , .after = Income) 
resident_profile %>%
    select(c(1:18))
```

### 3.4.1 Entries Check

To ensure data accuracy, the code chunk below checks the completeness of the data. Given that the data has a time period of one year, the code examines if the participants have entries for the entire time period.

```{r}
#| code-fold: true
#check if participants_id have entries for the entire year 
participant_counts <- resident_profile %>%
  group_by(participantId) %>%
  summarise(num_months = n_distinct(year_mth)) %>%
  ungroup()

filtered_count <- participant_counts %>%
  filter(num_months != 12) %>%
  nrow()

filtered_count 
```

It has been observed that there are **131 participants** who only have 1 entry on Mar 2022. To avoid inaccuracy, these group of participants will be excluded from the analysis.

### 3.4.2 Missing Values Check

Through the code chunk below, we confirmed that there are no missing values in resident_profile dataset.

```{r}
#Check for missing values
any(is.na(resident_profile))
```

### 3.4.3 Revised Resident's Profile Dataset

Given that we have removed duplicates in section 3.3.2, removed entries in section 3.4.1 and observed no missing values in section 3.4.2, the resident's profile data set have been revised. We will be using the `knitr: kable()` function to display the final dataset.

```{r}
#| code-fold: true
#create a revised dataframe to exclude id that do not have entries for the time period
resident_profile_rev <- resident_profile %>%
  group_by(participantId) %>%
  mutate(num_months = n_distinct(year_mth)) %>%
  ungroup() %>%
  filter(num_months == 12) %>%
  select(1:13,15:19)

#output for dataframe using knitr:: kable
kable(head(resident_profile_rev), "simple") 
```

# 4. Exploratory Data Visualization

In this section, we will design plots with interactivity for users to study and explore the data. The plots are created with the use of `giraph` , `plotly`, and `patchwork` packages.

## 4.1 Interactive Dashboard

A dashboard is created to provide an overview of the demographics of residents in City of Engagement across age group. Bar chart is chosen to show segments of information by comparing different categorical variables. A design layout is included in the code to better visualized the output through `patchwork`. Moreover, `tooltip` is used to highlight the specific age group at the point of the data.

```{r}
#| code-fold: true
#| fig-width: 12
#| fig-height: 6
#| warning : false

#Create new df by doing a full join with participants_counts
participants_new_rev <- full_join(
  participants_new, participant_counts, 
  by = c("participantId" = "participantId"))  %>%
  group_by(participantId) %>%
  #participants_counts contains the number of months the participants entries are in
  ungroup() %>%
  #filter away participantsId who do not have entries for the full year
  filter(num_months == 12) %>%
  select(-num_months)

#create tooltip to display age group 
participants_new_rev$tooltip <-c(paste0(
  "Age Group:", participants_new_rev$age_group))

#Bar chart for resident's age distribution
p1 <- ggplot(data= participants_new_rev,
      aes(x = age_group)) +
      geom_bar_interactive(aes(tooltip = participants_new_rev$tooltip, 
                               stackgroups = TRUE,
                               data_id= age_group)) + 
      scale_fill_manual(values = wes_palette("Chevalier1")) +
      xlab("Age Group") +
      ylab("No. of participants") +
      theme(axis.text.x=element_text(size=5)) +
      theme(axis.title.y=element_text(size=10)) +
      ylim(0,250)

#Bar chart for resident's household size distribution
p2 <- ggplot(data= participants_new_rev,
      aes(x = householdSize)) +
      geom_bar_interactive(aes(tooltip = participants_new_rev$tooltip, 
                               stackgroups = TRUE,
                               data_id= age_group)) +
      scale_fill_manual(values = wes_palette("Chevalier1")) +
      xlab("Size of Household") +
      ylab("") +
      theme(axis.text.x=element_text(size=5)) +
      theme(axis.title.y=element_text(size=10)) +
      ylim(0,350)


#Bar chart for resident's education level
p3 <- ggplot(data= participants_new_rev,
      aes(x = educationLevel)) +
      geom_bar_interactive(aes(tooltip = participants_new_rev$tooltip, 
                               stackgroups = TRUE,
                               data_id= age_group)) +
      xlab("Education Level") +
      ylab("") +
      theme(axis.text.x=element_text(size=5)) +
      theme(axis.title.y=element_text(size=10)) +
      ylim(0,350)

#Bar chart to visualize if residents have kids 
p4 <- ggplot(data= participants_new_rev,
      aes(x = haveKids)) +
      geom_bar_interactive(aes(tooltip = participants_new_rev$tooltip, 
                               stackgroups = TRUE,
                               data_id= age_group)) +
      ylab("") +
      theme(axis.text.x=element_text(size=5)) +
      theme(axis.title.y=element_text(size=10)) +
      ylim(0,400)

#Bar chart for residents' interest group 
p5 <- ggplot(data= participants_new_rev,
      aes(x = interestGroup)) +
      geom_bar_interactive(aes(tooltip = participants_new_rev$tooltip, 
                               stackgroups = TRUE,
                               data_id= age_group)) +
      xlab("Interest Group") +
      ylab("No. of participants") +
      theme(axis.text.x=element_text(size=5)) +
      theme(axis.title.y=element_text(size=10)) +
      ylim(0,120)

#Bar chart for residents' joviality in bins 
p6 <- ggplot(data= participants_new_rev,
      aes(x = joviality_bins)) +
      geom_bar_interactive(aes(tooltip = participants_new_rev$tooltip, 
                               stackgroups = TRUE,
                               data_id= age_group)) +
      xlab("Joviality") +
      ylab("No. of participants") +
      theme(axis.text.x=element_text(size=5)) +
      theme(axis.title.y=element_text(size=10)) +
      ylim(0,250)

#design layout for the patchwork figure
design <- "
  132
  132
  554
  554
  666
  666
"

girafe(code = print(p1 + p2 + p3 + p4 + p5  + p6 +
                      plot_layout(design = design,) + 
                      plot_annotation(title = 
                    "Demographics Insights of residents in City of Engagement",
                     theme = theme(plot.title = element_text(size = 20, hjust=0.5))
                      )), 
       width_svg = 12,
       height_svg = 6,
       options = list(
         opts_hover(css = "fill: #02401B;"),
         opts_hover_inv(css = "opacity:0.2;")
         )
       ) 
```

**Observations:**

-   The city is facing an aging population as age group is left-skewed.
-   Low interest for future education as education level is right-skewed.
-   Small family size with higher proportion of participants of not having Kids.
-   Uniformly distributed Interest Group. No preference among age group.
-   Joviality level seems to be decreasing at a decreasing rate

::: callout-note
## Interactivity

Click on the graph and hover around each demographics.

The respective age group will be displayed.
:::

## 4.2 Financial Health of Participants

To know more about the financial health of the participants, interactive `geom_point` is used to plot against the time period. `tooltip` is included to create a snapshot of the financial health status of the participants at the time period. In addition, the plot contains hover effect with the use of `data_id` aesthetic to highlight the trend of the participant's cash flow.

```{r}
#| code-fold: true
#| fig-width: 12
#| fig-height: 6

#tooltip output to display ID, Cashflow, Income, and Expenses
resident_profile_rev$tooltip <- paste0(
  "Participant's ID = ", resident_profile_rev$participantId,
  "\n Cashflow = ", format(resident_profile_rev$Cashflow, big.mark = ","),
  "\n Income = ", format(resident_profile_rev$Income, big.mark = ","),
  "\n Expenses = ", format(resident_profile_rev$Expenses, big.mark = ",")
                                  )
#tool_tip design
tooltip_css <- "background-color: lightgrey; #<<
font-style:bold; color: #446455;" #<<

ie <-  ggplot(data=resident_profile_rev) +
       geom_point_interactive(aes(x = year_mth, y = Cashflow,
                                   tooltip = resident_profile_rev$tooltip,
                                   data_id = participantId,
                                   #if Cashflow >0 = Green, else Red
                                   color = ifelse(Cashflow >= 0, 
                                                  "Above 0", "Below 0")
                                   )) +
        scale_color_manual(values = c("Above 0" = "#446455", 
                                      "Below 0" = "#C93312")) +
        #remove legend title 
      labs(color = "") +
      labs(title="Financial Health of Participants from Mar 2022 to Feb 2023") +
      ylab("Cashflow ($)") + xlab("Year-month") +  
      scale_y_continuous(labels = comma_format()) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
      theme_minimal()        |>

girafe(                                  
  ggobj = ie,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list( #<<
    opts_tooltip(css = tooltip_css), #<<
    opts_hover_inv(css = "opacity:0.1;") #<<
  )                                        
)   
   
```

**Observations:**

-   Majority of participants have positive cash flow (Good Credit Rating)
-   Negligible growth on cash flow
-   No drastic changes in participant's cash flow over time

## 4.3 The Circulation of Money

As illustrated earlier, the cash flow is a holding transaction. It is holded by the individual after repaying the necessary debt/expenses. To build more wealth, it is ideal for Wage to increase or for expenses to decrease.

With that in mind, a `line scatter pot` is used to identify the trend in various categories over a time period. The *participants* dataset have been revised to exclude participants who did not fulfill the entries check. It has been reformatted before executing a `full_join` with *participants_count* data set. The joined table concatenates the count of the year-mth period (num_months) that is not available in the dataset and filters it away.

The `plot_ly()` graph below shows the total amount circulated by the participants.

```{r}
#| warning: false
#| code-fold: true

#to reformat partcipantId, create year-mth, round off amount to 2.d.p
financial_journal_lessdup <- financial_journal_lessdup %>%
  mutate(participantId = as.character(participantId),
        year_mth = format(as.Date(financial_journal_lessdup$timestamp), "%Y-%m",
        amount = abs(round(amount,2)),
        .before = 1)) 

#Create new df by doing a full join with participants_counts
financial_journal_lessdup_lessentries <- full_join(
  financial_journal_lessdup, participant_counts, 
  by = c("participantId" = "participantId"))  %>%
  group_by(participantId) %>%
  #participants_counts contains the number of months the participants entries are in
  mutate(num_months = n_distinct(year_mth)) %>%
  ungroup() %>%
  #filter away participantsId who do not have entries for the full year
  filter(num_months == 12) %>%
  select(participantId, year_mth, category,amount)

#create another df to group by year_mth, category
grouped_data_rev <- financial_journal_lessdup_lessentries %>%
  #group the columns in the following order 
  group_by(year_mth, category) %>%
  summarize(total_amount = sum(amount)) 

#creating interactive graph
plot_ly(data = grouped_data_rev, 
  x = ~year_mth, y = ~total_amount, color = ~category,
  type = 'scatter', mode = 'line',
  hovertemplate = ~paste("Year-Month:", year_mth,
                         "<br>Amount:", format(total_amount, big.mark = ","))) |>

#Configure title and axes
  layout(title = "The Circulation of Money",
         xaxis = list(title = "Time Period"),
         yaxis = list(title = "Amount"))

```

**Observations:**

-   Income *(\~Wage)* is fairly constant except for Mar 2022. As Income varies with overtime, it can be inferred that the abnormally could be a form of bonus, grant, incentives given by the service centre
-   Expenses distribution seems to be constant with lesser spending on Education expenses . However, there is a slight fluctuation in Recreation expenses.

### 4.3.1 Spending Patterns of participants

As observed earlier, there are a slight fluctuation in Recreation expenses. Henceforth, Wage will be removed in the graph to better visualize the expenses.

`Plot_ly` graph is plotted to visualize the spending patterns.

```{r}
#| warning: false
#| code-fold: true

#create a new dataset
grouped_data_rev_new <- financial_journal_lessdup_lessentries %>%
  mutate(amount = abs(round(amount,2))) %>%
  #group the columns in the following order 
  group_by(category,year_mth) %>%
  summarize(total_amount = sum(amount)) 

# Filter out "Wage" category from the data frame
grouped_data_rev_newest <- grouped_data_rev_new %>% 
   filter(category != "Wage")

#creating interactive graph
plot_ly(data = grouped_data_rev_newest, 
  x = ~year_mth, y = ~total_amount, color = ~category,
  type = 'scatter', mode = 'line',
  hovertemplate = ~paste("Year-Month:", year_mth,
                         "<br>Amount:", format(total_amount, big.mark = ","))) |>

#Configure title and axes
  layout(title = "Total Expenses incurred by partcipants \nin City of Engagement from Mar 2022 - Feb 2023",
         xaxis = list(title = "Time Period"),
         yaxis = list(title = "Total Expenses"))

```

**Observations:**

-   Education remain constant throughout while Shelter decrease in Mar 2022 and remain constant. As identified, both expenses are a fixed expense.
-   Recreation expenses fluctuates more in comparison to Food expenses.

# 5. Confirmatory Data Analysis Visualization

In this section, we will focus more on the statistical testing that are used in Confirmatory Data analysis. The plots are created with the use of `ggbarstats` , `ggbetweenstats`, and `gscatterstats` packages.

## 5.1 Association Test between Age group and Education Level

As observed in Section 4.1 - Interactive Dashboard, we noticed that the age-group is left-skewed while the education level is right-skewed. Therefore, we would like to test if there is any association between the two variables. Notably, the association test is non-parametric and thus, does not have to conform to the normality assumption.

[At 95% confidence level,]{.underline}

Ho : No association exists between the age group and education level

H1: Association exists between the age group and education level

```{r}
#| code-fold: true
ggbarstats(data = resident_profile_rev, 
           x = educationLevel, y = age_group,
           xlab= "Age Group", ylab = "Education Level",
           title = "Comparison of Education level across Age Group",
           type = "nonparametric", conf.level = 0.95,
           package = "wesanderson", palette = "Chevalier1"
           )
```

From the graph, we observed that majority of the age group have a HighSchoolOrCollege education level whereas the interest for further study decreases as the age increase. Interestingly, young adults aged 20-29 have attained a higher education level than other age group.

From the test result above (p\<0.05) , we conclude that there is an association between the age group and education level as we **reject** the null hypothesis.

## 5.2 Differences in Joviality based on Education Level

As defined earlier, Joviality indicates the participant's overall happiness at the start of the study. We will like to found out if there is a difference in Joviality based on Education Level. Before testing our hypothesis, we will perform a normality assumption test at 95% confidence level.

### 4.4.1 Normality Assumption Test

Normal Quantile Plot (QQ Plot). For a normally distributed data values, the dot points should be with the 95% envelope or scatter very closely to the slope line in between the two envelopes lines. From the data visualization above, we can visually confirmed that the observed values failed to conform to the normality assumption.

At 95% confidence level:

Ho: the observed distribution resembles normal distribution

H1: the observed distribution failed to resemble normal distribution

We failed to reject the Ho if p-value is above the alpha value of 0.05.

We reject Ho if p-value is \<0.05.

Given that normality is not assumed, we will use the non-parametric test such as Wilcoxon test statistics.

```{r}
#| code-fold: true
#| fig-width: 12
#| fig-height: 6
qq_plot <- ggplot(data=resident_profile_rev,
                  aes(sample = joviality)) + 
                  stat_qq() +
                  stat_qq_line()
```

```{r}
ggplot(data=resident_profile_rev,
       aes(sample = joviality)) + 
  stat_qq() +
  stat_qq_line()
```

```{r}
#  ggplot(data = resident_profile_rev,
#         aes(x= joviality)) + 
#  geom_histogram(bins=10) +
#  xlab("Joviality") +
#  ylab("No.of Residents") +
#  geom_vline(aes(xintercept = average_joviality), col ="red", linewidth=1 ) +
#  annotate("text", x=0.37, y= 1400, label="Average Joviality:", size=4, color = "red") +    annotate("text", x=0.37, y= 1300, label=format(average_joviality, big.mark=","), size=4, color = "red") 
#  geom_vline(aes(xintercept = median_joviality), col ="green", linewidth=1 ) 

```

Based on the result above, we concluded that there is enough statistical evidence to reject the null hypothesis. Since the p-values fall below (p \< 0.05), normality is not assumed. Henceforth, we will use the Kruskal-Wallis test.

### 4.4.2 Kruskal-Wallis Test for Joviality across Education Level

We will test the following hypothesis at 95% Confidence Level:

Ho : the median Joviality across different education level is the same

H1: the median Joviality across different education level is not the same

```{r}
#| code-fold: true
#| fig-width: 12
#| fig-height: 6
ggbetweenstats(data = resident_profile_rev,
       x= educationLevel, y= joviality, type ="np",
       xlab= "Education Level", ylab = "Joviality",
       title = "Comparison of Joviality across Education Level",
       pairwise.comparisons = TRUE, pairwise.display ="ns", conf.level = 0.95,
       package = "wesanderson", palette = "Chevalier1"
       )
```

As seen above, the P-value is lower than the 0.05. As such, there is enough statistical evidence to reject the null hypothesis that the median joviality across education level is the same.

Additionally, we want to find out if there any distinct similarities between the district. Through the graph above, we discovered that not all pair comparison are statistically significant. The pair (Low and HighSchoolOrCollege) is not statistically significant with a P-value of 0.14, which is greater than 0.05. Thus, we cannot reject the null hypothesis that there is not differences between the joviality level between the pair.

## 4.5 Association between Age group and Joviality

We discovered that there is a statistical difference in joviality across Education Level, we would like to examine if the same applies and if there is an association between the age group. As Joviality is a continuous variable, we used `mutate` from `dplyr` to split it into 5-class variables, with each range at 20%.

[At 95% confidence level,]{.underline}

Ho : No association exists between the age group and joviality level

H1: Association exists between the age group and joviality level

```{r}
ggbarstats(data = resident_profile_rev, 
           x = joviality_bins, y = age_group,
          title = "Comparison of Joviality across age-group",
          pairwise.comparisons = TRUE, pairwise.display ="s", conf.level = 0.95,
          package = "wesanderson", palette = "Chevalier1"
           )
```

```{r}

```

## 5.2 Differences in joviality across income

We want to find out if there are any differences in joviality across income.

The following considerations have been made:

-   assume income increases over the years

-   opacity included to highlight the contrast

-   Tooltip to indicate variables

    -   participantID

    -   Income

    -   Cashflow

We will be using `plot_ly` for the interactive graph.

```{r}
plot_ly(data = resident_profile_rev, 
             x = ~joviality, y = ~Income,
        hovertemplate = ~paste("<br>Participant's ID:",participantId,
                               "<br>Cashflow:", Cashflow,
                              "<br>Income:", Income,
                               "<br>Expenses:", Expenses),
             
            type = "scatter",
            mode = "markers",
            marker = list(opacity = 0.7,sizemode = "diameter", 
                          line = list(width =0.1, color = "white"))) |>

#add title and labels to axis 
        layout(title = "Interactive scatterplot of Income vs Joviality" ,
         xaxis = list(title = "Joviality level") ,
         yaxis = list(title = "Income"))

```

**Observations:**

-   When income exceeds \$15k, joviality level decreases

-   it is similar to income range of \$10-\$15k where majority have low joviality

-   Joviality level are more spread out when income range below \$5k

::: callout-note
## Note

Happiness level of `joviality` ranges from 0 to 1. A score of 0.4 indicates that participant is dull while a score of 0.8 indicates joy.
:::

```{r}
DT::datatable(resident_profile_rev, class= "compact")
```

```{r}
d <- highlight_key(resident_profile_rev) 
p <- ggplot(d, 
            aes(Expenses, 
                Income)) + 
  geom_point(size =1) 

gg <- highlight(ggplotly(p),        
                "plotly_selected")  

dt <- DT::datatable(d,
                    selection =list(mode="single",
                                    target="column"))
crosstalk::bscols(gg,               
                  dt, 
                  widths = 6)        

```

```{r}
fj_cat <- financial_journal_lessdup %>%
  
#recode ID from dbl to chr, year_mth
    mutate(participantId = as.character(participantId),
         year = format(as.Date(financial_journal_lessdup$timestamp), "%Y"),
         mth = format(as.Date(financial_journal_lessdup$timestamp), "%m"),
         amount = abs(round(amount,2)),
         .before = 3) %>%
  
#group the columns in the following order 
  group_by(year, mth, category) %>%
  summarize(total_amount = sum(amount))

fj_cat
```

```{r}
ggplot(data = fj_cat,
       aes(x= category, y=total_amount, size = total_amount)) + 
         geom_point(alpha=0.8, show.legend= FALSE) +
         scale_size(range =c(2,12)) +
         labs(titles = "Year{as.integer(frame_time)}",
              x = "Category",
              y = "Total_amount") +
         transition_time(as.integer(year)) +
         ease_aes("linear") +
         scale_y_continuous(labels = comma_format()) 
```

## 5.3 Average Cashflow available for Residents across Age Group

```{r}
#| code-fold: true
#| fig-width: 12
#| fig-height: 6

tooltip_css <- "background-color:#C7B19C; #<<
font-style:bold; color:#446455;" #<<

tooltip <- function(y, ymax, accuracy = 1) {   #<<
  mean <- scales::number(y, accuracy = accuracy) #<<
  sem <- scales::number(ymax - y, accuracy = accuracy) #<<
  paste("Average Cashflow:", mean, "+/-", sem) #<<
} #<<

pp <- ggplot(data=resident_profile_rev, 
                   aes(x = age_group),
) +
  stat_summary(aes(y = Cashflow, 
                   tooltip = after_stat(  #<<
                     tooltip(y, ymax))),  #<<
    fun.data = "mean_se", 
    geom = GeomInteractiveCol,  #<<
    fill = "#D3DDDC"
  ) +
  stat_summary(aes(y = Cashflow),
    fun.data = mean_se,
    geom = "errorbar", width = 0.2, linewidth = 0.2
  ) +
  labs(title="Average Cashflow available to Residents by age_group") +
  ylab("Total Cashflow") +
  xlab("Age Group") 

girafe(ggobj = pp,
       width_svg = 8,
       height_svg = 8*0.618,
       options = list(    #<<
       opts_tooltip(    #<<
       css = tooltip_css)) #<<
       )
```

```{r}
ggplot(data =resident_profile_rev, 
       aes(x = Food, 
           y = age_group)) +
  geom_density_ridges(
    scale = 3,
    rel_min_height = 0.01,
    bandwidth = 3.4,
    fill = "#D3DDDC",
    color = "#446455"
  ) +
  scale_x_continuous(
    name = "English grades",
    expand = c(0, 0)
    ) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()
```

```{r}
ggscatterstats(
  data = resident_profile_rev,
  x = Expenses,
  y = Shelter,
  type = "nonparametric",
  marginal = TRUE,
  title = "Significant Test of Correlation between Expenses and Shelter",
  xlab = "Expenses",
  ylab = "Shelter"
)
```

```{r}
#plotting correlation between age and income 
ggscatterstats(data = resident_profile_rev,
  x = age, y = Income,
  type = "nonparametric",
  marginal = TRUE,
  title = "Significant Test of Correlation between Age and Income",
  xlab = "Age",
  ylab = "Income"
)
```

Cashflow / Income

```{r}
#| fig-width: 12
#| fig-height: 6
#| code-fold: true
#| warning: false

#plotting correlation between age and income across education level
edu_low <- ggscatterstats(data = resident_profile_rev |> 
                           filter(educationLevel == "Low"), 
                           x = Income, y = Cashflow,
                           type = "nonparametric") + 
           theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
           labs(title = "Low", 
           x = "Income", y = "Cashflow") +
           scale_y_continuous(labels = comma_format()) 

edu_hc <- ggscatterstats(data = resident_profile_rev |> 
                           filter(educationLevel == "HighSchoolOrCollege"), 
                           x = Income, y = Cashflow,
                           type = "nonparametric") + 
          theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
          labs(title = "High School", 
          x = "Income", y = "Cashflow") +
          scale_y_continuous(labels = comma_format()) 

edu_bach <- ggscatterstats(data = resident_profile_rev |> 
                           filter(educationLevel == "Bachelors"), 
                           x = Income, y = Cashflow,
                           type = "nonparametric") + 
            theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
            labs(title = "Bachelors", 
            x = "Income", y = "Cashflow") +
            scale_y_continuous(labels = comma_format()) 

edu_grad <- ggscatterstats(data = resident_profile_rev |> 
                           filter(educationLevel == "Graduate"), 
                           x = Income, y = Cashflow,
                           type = "nonparametric") + 
            theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
            labs(title = "Graduate", 
            x = "Income", y = "Cashflow") +
            scale_y_continuous(labels = comma_format()) 

#combined plot and ensure layout is in order
corr_edu <- ((edu_low + edu_hc) / (edu_bach + edu_grad) + plot_spacer())

#add labels
corr_edu + plot_annotation(title = "Correlation between Cashflow and Income", 
                           subtitle = "High correlation between Cashflow and Income at all education level",
                           theme = theme(
                             plot.title = element_text(size = 18),
                             plot.subtitle = element_text(size = 12)))

```

***Observations:***

-   Almost perfect positive linear relationship between Cashflow and Income (\>0.9)

-   As education level increases, the Sprearman correlation coefficient increases
