---
title: "In-class_Ex01"
author: "Oh Jia Wen"
---

## Getting started

1.  Using p_load() of pacman package to load tidyverse on

```{r}
pacman:: p_load(tidyverse) 
```

2.  Importing data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

## Plotting the graph 

### 1) Horizontal Bar Graph 

Changing the colors of plot panel background of theme_minimal() to light blue and the color of grid lines to white.

#### Output:

```{r}
ggplot(data= exam_data,
       aes(x = RACE)) +
       geom_bar() +
       coord_flip() +
       theme_minimal() +
       theme(panel.background = element_rect(fill = 'lightblue') ,
       panel.grid.minor=element_line(colour="white"),
       panel.grid.major=element_line(colour="white")) +
       ggtitle("Horizontal Bar Chart: \nLight Blue theme and White grid lines ") +  
       theme(plot.title = element_text(hjust = 0.5))
```
### 2) Vertical Bar Graph

With reference to the critics on the earlier slide, create a makeover looks similar to the figure on the right.

#### Output:

```{r}
ggplot(data= exam_data,
      aes(x = fct_infreq(RACE))) +
      geom_bar() +
      xlab("Race") +
      ylab("No.of\nPupils") +
      ylim(0,220) +
      geom_text(aes(label = paste(..count..,",", scales::percent(..count../sum(..count..),accuracy = 0.1))), 
      stat= "count", vjust = -0.5) +
      ggtitle("Vertical Bar Chart: \nSorted Frequency + Labelling ") +
      theme(plot.title = element_text(hjust = 0.5))
```
  

